---
title: "MP3"
author: "Fiona Miller, Starry Zhou, Morgan McLain-Smith"
date: "April 10, 2017"
output: html_document
---

## Problem Statement
We worked on project 3: genre multiplicity. To figure out the change of genre number of movies over the years 1900 to 2016, we visuzlized two changing trends: the change of the average genre number of each movie each year, and the change of the spread of movies with different numbers of genres each year.  

```{r setup, include=FALSE}
library(mdsr)
library(tidyverse)
library(RMySQL)
db <- dbConnect_scidb(dbname = "imdb")
```

## Average genre
## Total Films
```{r}
table2 <- db %>%
  dbGetQuery("SELECT id, production_year FROM title
             WHERE kind_id = 1;")
```

```{r}
table3 <- table2%>%
  select(production_year)%>%
  group_by(production_year) %>%
  summarise(num_films = sum(n()))
```

```{r}
table4 <- table3 %>%
  mutate(total=sum(num_films))
```

## Total Genres
```{r}
movieGenreAndYear <- db %>%
  dbGetQuery("SELECT mi.movie_id, mi.info, t.production_year 
              FROM movie_info mi
              INNER JOIN title t ON t.id = mi.movie_id
              WHERE mi.info_type_id = 3  
                    AND t.kind_id = 1;") 
```

```{r}
t3 <- movieGenreAndYear %>%
  group_by(production_year) %>%
  summarise(num_genre = n())
```

## Average number of Genres
```{r}
g1 <- t3 %>%
  inner_join(table3, by = "production_year")
```

```{r}
g2 <- g1 %>%
  mutate(avg_genre = num_genre/num_films)
```
```{r}
g3<- g2%>%
  filter(production_year>=1900,production_year<=2016)
```

## Visualization
```{r}
vis <- ggplot(g3, aes(x=production_year,y=avg_genre))+
geom_line(aes(fill = avg_genre))
vis
```

## The spread of movies with different numbers of genres each year

```{r}
numGenrePerMovieByYear <- movieGenreAndYear%>%
  group_by(movie_id, production_year) %>%
  summarize(genresPerMovie = n()) %>%
  ungroup() %>%
  group_by(production_year, genresPerMovie) %>%
  summarize(countMovies = n()) %>%
  filter(production_year<=2016, production_year>=1900)%>%
  mutate(genresPerMovie = factor(genresPerMovie, levels = 15:1))


```

```{r}
ggplot( numGenrePerMovieByYear, aes(x= production_year, y = countMovies, fill = genresPerMovie))+
  geom_bar(position = "fill", stat="identity")+
  geom_vline(xintercept = 1947, linetype=2)
```

## Explanation
We expected to see some major changes based on influential events in film field or after the production of some successful movies. From the two graphs, it seems that before the year of 1987, the genre numbers of movies gradually decreased for a short while and then change to a different direction. From 1987 to 2014 there is a constant increase, but after the genre number reaches the peak of average genre number per movie (average genre number per movie: 2.36) in 2014, it decreased again. Therefore, it seems that although clearly, the average numbers of genres per movie generally over the years, the overall change of genre numbers per movie is gradual and unpredictable. 